Flexibility of interactions depends on habitat suitability

### 2019-06-14 paper/manuscript.Rmd
Continue methods of species cleaning

---
title: "Flexibility of interactions depends on habitat suitability"
author: 
  - "Audrey Lustig^2^ ([audrey.lustig@canterbury.ac.nz](audrey.lustig@canterbury.ac.nz))"
  - "Jason M. Tylianakis^1^ ([jason.tylianakis@canterbury.ac.nz](jason.tylianakis@canterbury.ac.nz))"
  - "Daniel B. Stouffer^1^ ([daniel.stouffer@canterbury.ac.nz](daniel.stouffer@canterbury.ac.nz))"
  - "E. Fernando Cagua^1^ ([efc29@uclive.ac.nz](mailto:efc29@uclive.ac.nz))"
bibliography: 
  - bibliography.bib
  - int-bibliography.bib
csl: pnas.csl
output: #word_document
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
fontsize: 11pt
# classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[S-]{supp-info}
  - \usepackage[export]{adjustbox}
editor_options: 
  chunk_output_type: console
---

@@

@@

all together we found XX instances of species that were included with a datab

>>>>

---
title: "Flexibility of interactions depends on habitat suitability"
author: 
  - "E. Fernando Cagua^1^ ([efc29@uclive.ac.nz](mailto:efc29@uclive.ac.nz))"
  - "Audrey Lustig^2^ ([audrey.lustig@canterbury.ac.nz](audrey.lustig@canterbury.ac.nz))"
  - "Jason M. Tylianakis^1^ ([jason.tylianakis@canterbury.ac.nz](jason.tylianakis@canterbury.ac.nz))"
  - "Daniel B. Stouffer^1^ ([daniel.stouffer@canterbury.ac.nz](daniel.stouffer@canterbury.ac.nz))"
bibliography: 
  - bibliography.bib
  - int-bibliography.bib
csl: pnas.csl
output: #word_document
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
fontsize: 11pt
# classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[S-]{supp-info}
  - \usepackage[export]{adjustbox}
editor_options: 
  chunk_output_type: console
---

@@

```{r scientific-names-cleaning-info}
loadd(checked_sp_names, spp)

checked_sp_names <- checked_sp_names %>%
  # remove variant and subspecies abbreviations from the retrieved species
  # names, it is clear that they are one when there is 3 words or more in there
  mutate(sp_name = stringr::str_replace_all(sp_name,"var. ", ""),
         sp_name = stringr::str_replace_all(sp_name,"ssp. ", ""),
         sp_name_in_gnr = !is.na(gnr_score), 
         sp_name_itis_invalid = itis %in% c("invalid", "not accepted"), 
         # good species name has been found and its not invalid
         good_sp_name = sp_name_in_gnr & ! sp_name_itis_invalid, 
         good_queried_sp_name = good_sp_name & sp_name == queried_sp_name, 
         sp_name_rank = get_name_rank(sp_name), 
         queried_sp_name_rank = get_name_rank(queried_sp_name)) %>%
  group_by(queried_sp_name) %>%
  mutate(corrected_typo = any(gnr_score == 0.750) & all(sp_name_rank == queried_sp_name_rank),
         corrected_typo = tidyr::replace_na(corrected_typo, FALSE),
         good_queried_sp_name = any(good_queried_sp_name), 
         n_spellings = n_distinct(sp_name), 
         any_sp_itis_invalid = any(sp_name_itis_invalid), 
         alt_sp_name_found = n_spellings > 1 & any_sp_itis_invalid) %>%
  ungroup()

n_sp_names <- n_distinct(spp$sp_name)

n_sp_names_sp_level <- filter(spp, !sp_unidentified) %$% 
  n_distinct(sp_name)

n_sp_names_gen_level <- filter(spp, sp_unidentified, !gen_unidentified) %$% 
  n_distinct(sp_name)

n_no_tax_info <- filter(spp, gen_unidentified) %$% n_distinct(sp_name)

n_sp_correct_in_data <- checked_sp_names %>%
  filter(good_queried_sp_name) %$%
  n_distinct(queried_sp_name)

n_sp_corrected_typos <- checked_sp_names %>%
  filter(corrected_typo) %$% 
  n_distinct(queried_sp_name)

sp_with_synonyms <- checked_sp_names %>%
  filter(alt_sp_name_found)

n_sp_with_synonyms <- sp_with_synonyms %$%
  n_distinct(queried_sp_name)

n_sp_not_found <- checked_sp_names %>%
  filter(!good_queried_sp_name) %>% 
  filter(!alt_sp_name_found, !corrected_typo) %$%
  n_distinct(queried_sp_name)

checked_sp_names %>%
  filter(!good_queried_sp_name, !alt_sp_name_found) %$% n_distinct(queried_sp_name)
```

@@

All together, before correcting typos and synonyms, the data sources contained `r format(n_sp_names, big.mark = ",")` unique interacting species. 
From that total, `r format(n_sp_names_sp_level, big.mark = ",")` were originally identified to the species-level, `r format(n_sp_names_gen_level, big.mark = ",")` to the genus level, and no taxonomic information was avaliable for the remaining `r format(n_no_tax_info, big.mark = ",")` organims. 
We were able to confirm the validity of `r format(n_sp_correct_in_data, big.mark = ",")` of the scientific names used to identify organisms to the species level (roughly `r scales::percent(n_sp_correct_in_data/n_sp_names_sp_level, 1)`). 
We assessed the validity of a scientific name by querying the Global Names Resolver database (https://resolver.globalnames.org) which includes taxonomic data from `r dplyr::n_distinct(taxize::gnr_datasources()$title)` taxonomic sources. 
Specifically, we used the function `gnr_resolve` from the R package `taxize `r as.character(packageVersion('taxize'))`` [@chamberlain_taxize_2013; @chamberlain_taxize_2019].

From the remaining `r format(n_sp_names_sp_level - n_sp_correct_in_data, big.mark = ",")` names, we were able to identify and the correct missspelling of `r format(n_sp_corrected_typos, big.mark = ",")` organism names. 
Furthermore we found alternative names for `r format(n_sp_with_synonyms, big.mark = ",")` species. 
Most commonly, due to the name being an invalid or not accepted synonym of a species, subspecies, or plant variety (\@ref(tab:tab-sp-synonym-reasons)). 
We were unable to automatically correct `r format(n_sp_not_found, big.mark = ",")` names in the databases.

```{r tab-sp-synonym-reasons}
sp_with_synonyms %>%
  filter(!is.na(itis_reason)) %>%
  group_by(queried_sp_name) %>%
  summarise(itis_reason = first(itis_reason)) %>%
  group_by(itis_reason) %>%
  count(sort = T) %>%
  knitr::kable(col.names = c("reason of species alternative name", "n"), 
               caption = "tablesita", 
               booktabs = TRUE) %>%
  kableExtra::kable_styling(font_size = 8) 
```
